Shapes analysis for time series.

### 2024-05-22 NEURIPS24/appendix.tex
experiment all
appendix 0

\section{Tuning the hyperparameters of the TS-LDDMM kernel given in \eqref{eq:kernel_TAS}}
\label{appendix:kernel_TS_LDDMM}
The parameter $\sigma_{T,0}$ should be chosen \textit{large} compared the sampling frequency $f_e$ and compared to average standard deviation $\bar{\sigma}_s$ of the time series, e.g $\sigma_{T,0}=100$ as $\bar{\sigma}_s\approx f_e\approx  1 $.
It makes the time transformation smoother. If $\sigma_{T,0}$ is too small, for instance, $\sigma_{T,0}=f_e $, the effect of the time deformation is too localized, and there are not enough samples to make it visible.

@@

\subsection{Optimization details of TS-LDDMM~\eqref{eq:general_optimization_problem}}
\label{appendix:optimizers_details}

@@

\subsection{Identifiability experiment}
This experiement only involves the TS-LDDMM method in two different settings: 
\begin{itemize}
  \item \textbf{The velocity field kernel $K_G$ is well-specified:} The velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,100,1,1)$, the varifold loss kernels $(k_{pos},k_{dir})$ are set to $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}) = (2,1,2,0.6)$, and the optimizer has 400 steps with a maximum stepsize $\eta_M$ of 0.05.
  \item \textbf{The velocity field kernel $K_G$ is missspecified:} The velocity field kernel $K_G$ is set with  $(c_0,c_1,\sigma_{T,1}) = (1,0.1,1)$, $\sigma_{T,0}$ ranging in $(1,5,10,50,100,200,300)$, and $\sigma_x$  ranging in $(0.1,1,10,100)$. The varifold loss kernels $(k_{pos},k_{dir})$ are set to $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}) = (2,1,2,0.6)$, and the optimizer has 400 steps with a maximum stepsize $\eta_M$ of 0.05.
\end{itemize}
TS-LDDMM parameters of the experiment are summarized in \Cref{appendix:table:set_up_exp_practical}

\subsection{Mice respiratory behaviors experiment}

@@

\paragraph{LDDMMM parameters.} Note that varifold losses are unchanged between TS-LDDMM and LDDMM. Compared to TS-LDDMM, the convergence of LDDMM is more sensitive to the maximum stepsize $\eta_m$, which must remain small for LDDMM to guarantee the convergence.
\begin{itemize}
  \item \textbf{Before exposure:} The velocity field kernel $K_G$ is an anysotropic Gaussian kernel with parameters $\sigma_{T} =150$ for the time dimension and $\sigma_x = 2$ for space dimensions. 
  The varifold loss is the sum of three varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3): $\left((5,2,5,1),(2,1,2,0.6),(1,0.6,1,0.6)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$. The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.01.
  \item \textbf{Before/after exposure:} The velocity field kernel $K_G$ is an anysotropic Gaussian kernel with parameters $\sigma_{T} =220$ for the time dimension and $\sigma_x = 2$ for space dimensions.
  The varifold loss is the sum of four varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3, Varifold 4): $\left((30,2,30,1),(5,2,5,1),(2,1,2,0.6),(1,0.1,1,0.1)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$. The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.01.
\end{itemize}
\subsection{Benchmark}

A report of all the hyperparameters selected is given in \Cref{appendix:table:set_up_exp_practical}.

>>>>

\section{Tuning the hyperparameters of the TS-LDDMM velocity field kernel}
\label{appendix:kernel_TS_LDDMM}
The parameter $\sigma_{T,0}$ should be chosen \textit{large} compared the sampling frequency $f_e$ and compared to average standard deviation $\bar{\sigma}_s$ of the time series, e.g $\sigma_{T,0}=100$ as $\bar{\sigma}_s\approx f_e\approx  1 $.
It makes the time transformation smoother. If $\sigma_{T,0}$ is too small, for instance, $\sigma_{T,0}=f_e $, the effect of the time deformation is too localized, and there are not enough samples to make it visible.

@@

\subsection{Optimization details of TS-LDDMM \& LDDMM}
\label{appendix:optimizers_details}

@@

\subsection{Settings for experiment: analysis of respiratory behavior in mice}
\label{appendix: mice_exp_setting}

@@

\paragraph{LDDMMM parameters.} Note that varifold losses are unchanged between TS-LDDMM and LDDMM. Compared to TS-LDDMM, the convergence of LDDMM is more sensitive to the maximum stepsize $\eta_m$, which must remain small for LDDMM to guarantee the convergence.
\begin{itemize}
  \item \textbf{Before exposure:} The velocity field kernel $K_G$ is an anysotropic Gaussian kernel with parameters $\sigma_{T} =150$ for the time dimension and $\sigma_x = 2$ for space dimensions. 
  The varifold loss is the sum of three varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3): $\left((5,2,5,1),(2,1,2,0.6),(1,0.6,1,0.6)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$. The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.01.
  \item \textbf{Before/after exposure:} The velocity field kernel $K_G$ is an anysotropic Gaussian kernel with parameters $\sigma_{T} =220$ for the time dimension and $\sigma_x = 2$ for space dimensions.
  The varifold loss is the sum of four varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3, Varifold 4): $\left((30,2,30,1),(5,2,5,1),(2,1,2,0.6),(1,0.1,1,0.1)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$. The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.01.
\end{itemize}

\subsection{Settings for experiment: TS-LDDMM representation identifiability}
\label{appendix: settings_identifiability}
This experiment only involves the TS-LDDMM method in two different settings: 
\begin{itemize}
  \item \textbf{The velocity field kernel $K_G$ is well-specified:} The velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,100,1,1)$, the varifold loss kernels $(k_{pos},k_{dir})$ are set to $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}) = (2,1,2,0.6)$, and the optimizer has 400 steps with a maximum stepsize $\eta_M$ of 0.05.
  \item \textbf{The velocity field kernel $K_G$ is missspecified:} The velocity field kernel $K_G$ is set with  $(c_0,c_1,\sigma_{T,1}) = (1,0.1,1)$, $\sigma_{T,0}$ ranging in $(1,5,10,50,100,200,300)$, and $\sigma_x$  ranging in $(0.1,1,10,100)$. The varifold loss kernels $(k_{pos},k_{dir})$ are set to $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}) = (2,1,2,0.6)$, and the optimizer has 400 steps with a maximum stepsize $\eta_M$ of 0.05.
\end{itemize}

\subsection{Settings for experiment: Robustness to irregular sampling}
This experiment is inspired from \cite{oh2024stable} where authors performs an extensive comparison of Neural Ordinary Differential Equations (Neural ODEs) methods~\cite{kidger2020neural}. Methods, network architectures, and training strategy are taken from its associated Github\footnote{\url{https://github.com/yongkyung-oh/Stable-Neural-SDEs}} and describe in what follows.
\label{appendix: settings_robustness}
\subsubsection{Benchmark methods}
In related work, we give an overview of Neurals ODEs methods and its relation with TS-LDDMM.  
\begin{itemize}
  \item RNN-based methods: Baseline reccurent neural networks including RNN~\cite{medsker1999recurrent}, LSTM~\cite{hochreiter1997long}, and GRU~\cite{chung2014empirical}.

  \item Attention-based methods: Multi-Time Attention Networks (MTAN)~\cite{shukla2021multi} and Multi-Integration Attention Module (MIAM)~\cite{lee2022multi}. Both handles multivariate time series irregularly sampled with attention mechanisms.
  \item Neural ODEs: ODE-LSTM~\cite{lechner2020learning} a form of Neural-ODEs used to learned continous latent representations. 
  \item Neural SDEs:  Neural SDE~\cite{liu2019neural} and Neural LNSDE~\cite{oh2024stable} have been proposed to model randomness in time-series using drift and diffusion terms as an extension of Neural-ODEs. 
  \item Shape-Analysis methods: TS-LDDMM (ours) and LDDMM~\cite{glaunes2008large}. From shape analysis, both methods learn representations by solving ODEs parametrized with Kernels. While both methods hamdle multivariate signal irregularly sampled, TS-LDDMM is specifically desgined for time series.
\end{itemize}

\subsection{Settings for experiment: Classification benchmark on regularly sampled datasets}
\label{appendix: settings_classification}
